Dynamics of international giving: Exploring individual donor preferences

### 2020-06-18 manuscript/manuscript.Rmd
Severe cutting down pre Suparna

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private funding flows. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO regulatory environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We conclude by discussing the implications of our results on the scholarship and practice of nonprofits using different framing appeals to increase fundraising at a time when traditional funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "FILL"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon commonly known as "closing civic space" [@Carothers:2015; @DPR:2015]. Repressive governments have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which typically rely on funding from donors across the world. This is particularly the case in countries with weak institutions and poor governance, where international donors typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. As a result of this global crackdown on civil society, however, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to engage in philanthropy.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to $427.71 billion in 2018, 85.7% of which was through individuals [givingusa:2019]. Much of the existing research on donations to international causes is restricted to elite, high net-worth donors [@USTrust2014], but these large amounts demonstrate that it is imperative to understand the motivations and attributes of more ordinary individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about regulatory environments abroad, as well as organizational-level attributes of NGOS— such as their main issues of focus and sources of funding—affect donors' willingness to donate internationally?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974; @CrosonShang:2011; @Tremblay-BoirePrakash:2017], We leverage an online survey experiment using Amazon's Mechanical Turk to assess how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: NGO issue area and funding sources. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing stricter regulatory environments.

@@

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how these frames interact with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though our focus in this paper is one private funding, it is essential to differentiate between the two sources. Public aid (more commonly known as traditional donor aid) channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds—which are comparatively newer international donors compared to traditional donor aid—can come from foundations, private corporations, and individuals.

In this paper, we specifically look at individual-level determinants to international philanthropy. Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international charities, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@brainard2007; @Carothers:2015] and public aid decisions—especially about final recipients and amounts—are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. For instance, unlike established donors like the United States and other OECD countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

@@

Understanding the motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. Prior research has demonstrated that restrictions on cross-border funding has reduced official aid and official development assistance to repressive countries [@citeme]. But data from the Foundation Center shows no correlation between amount of funding flows from US private foundations to a given country and that country's score on the Index of Philanthropic Freedom. Additionally, preliminary evidence shows that "US foundations remain committed to channeling funds to countries where the legal environment is growing increasingly unfriendly." [@citeme-too]. As such, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to countries with restrictive legal environments.

## Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) the role of social capital and social networks; (2) the role of individual experiences such as higher education and levels of religiosity; and (3) third party certifications.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature concentrates on "the networks of formal and informal relationships to which people are associated," [@havens2007, 240] or associational capital. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations, since volunteering exposed individuals to the need for charitable donations and made philanthropy more compelling [@hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable giving.

@@

Finally, information about organizational characteristics and third party certifications can also impact levels of individual philanthropy, through results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

While this robust literature looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This domestic focus is warranted, though: on average, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, this focus on domestic philanthropy causes means that that the precise causal mechanisms through which cross-border charity can be motivated and mobilized is still under-examined [@desaikharas, 517]. This problem is further compounded by the fact that there is greater paucity of data and diversity of sources regarding international philanthropy. Indeed, most research on individual-giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014].[^highnetworth] Given the increasing importance of more ordinary individual donations for international NGOs, it is crucial to better understand what motivates private donors to engage in internationally-focused philanthropy.

As explored previously, individual private donors do not necessarily mirror the choices made by public donors. Individuals rely on signals and heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals and heuristics, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may be therefore more reliant on framing. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). We made two minor deviations from the original preregistration, described in detail in the online appendix.[^deviations] The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:[^consort]

@@

Without considering issue or funding, we find that respondents are not more likely to donate to legally besieged INGOs (H~1a~). As seen in @fig:likely-diffs(A), those exposed to the crackdown condition tend to have a slightly higher probability of donating (`r percent_1(filter(likely, title == "Crackdown")$theta_2)`) than those in the control condition (`r percent_1(filter(likely, title == "Crackdown")$theta_1)`), but the difference is not significant (`r filter(likely, title == "Crackdown")$delta`; `r filter(likely, title == "Crackdown")$p_neq0`). We find similar results when looking at the isolated effects of issue area and funding. Both differences are positive, but the posterior probability of being larger than zero is low (`r filter(likely, title == "Issue")$delta`; `r filter(likely, title == "Issue")$p_neq0`; `r filter(likely, title == "Funding")$delta`; `r filter(likely, title == "Funding")$p_neq0`). We thus do not initially find evidence that crackdowns (or issue or funding) alone increase the likelihood of donating.

@@

While the results from the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are a few caveats to consider when interpreting these results, generalizing them to other contexts, and applying them to nonprofit fundraising. First, the experiment measures preferences (willingness to donate) rather than behavior (actual donations). While the former may not always neatly translate into the latter, there is evidence that individuals’ willingness to donate (as measured in the experiment) translates well into behavior. The wording of the survey experiment uses an informational frame, where the focus is on educating respondents, rather simply presenting them with facts. Prior research has shown that informational frames—as opposed to the use of personal frames, which draw the audience’s attention to the plight of an individual, or motivational frames, which are meant to motivate individuals to act by creating feels of agency and efficacy—are important in generating donations [@McEntireLeibyKrain:2015].[^frametypes]

Moreover, prior research has found that informational frames are a necessary precondition for encouraging action in human rights nonprofit campaigns, and individuals are more likely to donate and undertake other actions when these informational frames are combined with personal frames [@McEntireLeibyKrain:2015].[^personalframes] This suggests that if governments attempt to repress international human rights nonprofits, then these organizations can potentially increase their resource flows by using a combination of these two frames. The survey instrument used above shows substantive increase in donors' willingness to donate to human rights groups facing crackdown, especially those that are privately-funded, and this result emerges from the use of just informational frames. If combined with personal frames, research indicates these donations may further increase.

It is also worth noting that there is evidence, at least domestically, that restrictions on nonprofit activities spur changes in private donors’ actual giving patterns (and not merely their preferences). For instance, following the 2016 U.S. presidential election, after Texas pledged to remove Planned Parenthood from its Medicaid program and a number of other states proposed new restrictions on their work, donations to the organization increased by over 40 times the usual amount [@citeme-as-well].

@@

Dozens of countries have sought to restrict NGO funding and programming over the past few decades. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities—especially international programming—is costly and time-consuming, donors use heuristics when deciding which organizations to support. Existing research has shown that donors take cues and heuristics from their social networks and from organizational characteristics like GuideStar ratings. In this paper, we argue that other informational cues can influence donor preferences. We find that the domestic political environments of NGO host countries can indeed serve as one such heuristic. Our results show that while crackdowns do not consistently directly influence individual preferences of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other organizational heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately funded human rights NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This article should be viewed as an exploratory study. While the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are limits to its external validity. Our sample captures people who are relatively more likely to donate to charities and thus does not reflect the general U.S. population. Our findings do not apply to the total population—we still know very little about how anti-NGO crackdowns change perceptions of nonprofits in general. Future research should look to testing the generalizability of our findings to individual donors in other countries, as well as determining the extent to which these preferences translate into behavior. For instance, the nonprofit sector in the U.S. relies much more heavily on government funding than in Europe, where large organizations like Amnesty International and Médecins Sans Frontières receive very little support from their home governments [@Stroup:2012], which means that donors in these countries will have different preferences when choosing to donate to government- and privately-funded organizations. Future research should also look more narrowly at donors who regularly give to international causes, as these individuals are likely more aware of poltiical trends in the countries where their preferred organizations work.

@@

[^deviations]: First, in the text of our preregistration, we inadvertently only specified non-nested versions of our hypotheses, but in our analysis plan, we describe the nested hypotheses. Second, in our analysis plan we stated that we would test for statisitcal differences using regression models with increasing numbers of interaction terms. This became unwiedly when interpreting results, though, and we instead simplified the analysis by estimating group means directly. The estimates are identical.

[^consort]: The CONSORT diagram in the online appendix shows the balance of respondents across different vignette combinations.

@@

[^internationalaffairs]: Giving USA defines the sector of international affairs to include any nonprofit organization working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs. This classification is based upon the National Taxonomy of Exempt Entities Core which codes groups based upon their field of interest.

[^exceptions]: For exceptions, see @Micklewright:2009 on UK, @Rajan:2009 on Canada, @wiepking2010 on Netherlands, @CasaleBaumann:2015 on the USA.

[^highnetworth]: High-net worth households are defined as having greater than $200,000 in annual household income and/or with more than $1,000,000 in assets, excluding primary residence.

>>>>

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private funding flows. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO regulatory environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We conclude by discussing the implications of our results on the scholarship and practice of nonprofits using different framing appeals to increase fundraising at a time when traditional funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "8,031"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

@@

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon known as "closing civic space" [@Carothers:2015; @DPR:2015]. Repressive governments have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which typically rely on funding from donors across the world. This is particularly the case in countries with weak institutions and poor governance, where international donors typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. As a result of this global crackdown on civil society, however, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to engage in philanthropy.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to \$427.71 billion in 2018, 85.7% of which was through individuals [@givingusa:2019]. Much of the existing research on donations to international causes is restricted to elite, high net-worth donors [@USTrust2014], but these large amounts demonstrate that it is imperative to understand the motivations and attributes of more ordinary individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about repressive regulatory environments abroad affect donors' willingness to donate internationally, and how does knowledge of legal crackdown interact with other organizational attributes?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. We use an survey experiment to assess how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: issue area and funding source. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing stricter regulatory environments.

@@

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard [@ChongDruckman:2007]. However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how these frames interact with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

@@

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though we focus on private funding in this article, it is essential to differentiate between the two sources. Public aid, or more traditional donor aid, channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds come from foundations, private corporations, and individuals. We are specifically interested in individual-level determinants to private international philanthropy. Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international charities, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@Carothers:2015] and public aid decisions—especially regarding final recipients and amounts—are not easily accessible. There is also heterogeneity in how donor governments allocate funds. For instance, unlike established donors like the United States and other OECD countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

@@

Understanding the motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. While anti-NGO restrictions have reduced official aid to repressive countries [@DupuyPrakash:2018], private foundations have continued to channel funds to countries with unfriendly legal environments [@Foundation-Center:2018]. Accordingly, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to organizations facing legal crackdown abroad.

# Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) the role of social and associational capital, (2) the role of individual experiences such as higher education and levels of religiosity, and (3) third-party certifications.

Prior research has found that those who participate in a variety of associations and build social capital are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations, since volunteering exposed individuals to the need for charitable donations and made philanthropy more compelling [@hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable giving.

@@

Finally, information about organizational characteristics and third-party certifications can also impact levels of individual philanthropy, through results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

While this robust literature looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This domestic focus is warranted, though: on average, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. Accordingly, there is a lack of data and theory regarding international philanthropy. Most research on individual donors to international causes has been restricted to elite, high net-worth donors (those with more than $200,000 in annual income or $1 million in assets) [@USTrust2014]. Given the increasing importance of more ordinary individual donations for international NGOs, it is crucial to better understand what motivates private donors to engage in internationally-focused philanthropy.

Individuals rely on heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may therefore rely framing to control perceptions. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different frames donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, CONSORT diagram, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

@@

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because it fits within each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:

@@

In isolation we find that respondents are not more likely to donate to legally besieged INGOs (H~1a~). As seen in @fig:likely-diffs(A), those exposed to the crackdown condition tend to have a slightly higher probability of donating (`r percent_1(filter(likely, title == "Crackdown")$theta_2)`) than those in the control condition (`r percent_1(filter(likely, title == "Crackdown")$theta_1)`), but the difference is not significant (`r filter(likely, title == "Crackdown")$delta`; `r filter(likely, title == "Crackdown")$p_neq0`). We find similar results when looking at the isolated effects of issue area and funding. Both differences are positive, but the posterior probability of being larger than zero is low (`r filter(likely, title == "Issue")$delta`; `r filter(likely, title == "Issue")$p_neq0`; `r filter(likely, title == "Funding")$delta`; `r filter(likely, title == "Funding")$p_neq0`). We thus do not initially find evidence that crackdowns (or issue or funding) alone increase the likelihood of donating.

@@

While the results from the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are a few caveats to consider when interpreting these results, generalizing them to other contexts, and applying them to nonprofit fundraising. First, the experiment measures preferences (willingness to donate) rather than behavior (actual donations). While the former may not always neatly translate into the latter, there is evidence that individuals’ willingness to donate (as measured in the experiment) translates well into behavior. The wording of the survey experiment uses an informational frame, where the focus is on educating respondents, rather simply presenting them with facts. Prior research has shown that informational frames—as opposed to the use of personal frames, which draw the audience’s attention to the plight of an individual, or motivational frames, which are meant to motivate individuals to act by creating feels of agency and efficacy—are important in generating donations [@McEntireLeibyKrain:2015].

Moreover, prior research has found that informational frames are a necessary precondition for encouraging action in human rights nonprofit campaigns, and individuals are more likely to donate and undertake other actions when these informational frames are combined with personal frames [@McEntireLeibyKrain:2015].[^personalframes] This suggests that if governments attempt to repress international human rights nonprofits, these organizations can potentially increase donations by combining these two frames. Our experiment shows a substantial increase in donors' willingness to donate to human rights groups facing crackdown, especially those that are privately-funded, and this result emerges from the use of only informational frames. If combined with personal frames, research suggests these donations may further increase.

It is also worth noting that there is evidence, at least domestically, that restrictions on nonprofit activities spur changes in private donors' actual giving patterns. For instance, after Texas and other states pledged to impose new restrictions on Planned Parenthood following the 2016 election, donations to the organization increased more than 40 times the usual amount [@Cooney:2016].

@@

In the era of closing civic space, dozens of countries restricted international NGO funding and programming. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many INGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities—especially international programming—is costly and time-consuming, donors use heuristics when deciding which organizations to support. Existing research has shown that donors take cues from their social networks and from organizational characteristics. In this paper, we argue that other informational heuristics can influence donor preferences. We find that the domestic political environments of INGO host countries can indeed serve as one such signal. Our results show that while crackdowns do not consistently directly influence individual preferences of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other organizational heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately funded human rights NGOs.

This article should be viewed as an exploratory study. While the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are limits to its external validity. Our sample captures people who are relatively more likely to donate to charities and thus does not reflect the general U.S. population. Our findings do not apply to the total population—we still know very little about how anti-NGO crackdowns change perceptions of nonprofits in general. Future research should look to testing the generalizability of our findings to individual donors in other countries, as well as determining the extent to which these preferences translate into behavior. For instance, the nonprofit sector in the U.S. relies much more heavily on government funding than in Europe, where large organizations like Amnesty International and Médecins Sans Frontières receive very little support from their home governments [@Stroup:2012], which means that donors in these countries will have different preferences when choosing to donate to government- and privately-funded organizations. Future research should also look more narrowly at donors who regularly give to international causes, as these individuals are likely more aware of political trends in the countries where their preferred organizations work.

@@

@@

[^internationalaffairs]: According to Giving USA, international affairs includes nonprofits working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs.

[^exceptions]: For exceptions, see @Micklewright:2009 on the UK, @Rajan:2009 on Canada, @wiepking2010 on the Netherlands, and @CasaleBaumann:2015 on the USA.

[^personalframes]: Personal frames humanize victims and entice action by invoking an emotional reaction. For instance. people are more be generous when donating to a human rights campaign with a single victim, rather than a statistical group of victims [@Slovic:2007].

[^NGO]: We use the terms nonprofit organizations and NGOs interchangeably. We consider international NGOs to be nongovernmental nonprofit organizations with members from two or more countries and that work in countries outside of their home country.
